---
title: "Lab4_LilRascals"
authors: "Matthew Copeland, Kelsey McKenna, James Burke, Shreeya Basrur"
date: "February 7, 2019"
output: html_document
---

```{r}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`) #renaming `bwt/oz` because the "/" sign requires special care
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Team Plots
```{r}
ggplot(data = babies)+
  geom_point(aes(x = id, y = date))
```
## What is the relationship between babies weight and whether the mother smokes accounting for a mother's weight during the pregnancy. 
* This is an interesting question because the answer could help us to find a better understand of how smoking impacts one's children. This is important because knowing this may help to discourage people from continuing to smoke or starting smoking, and encourage those who want to improve their babies well-being. Especially, in regards to mothers and how it impacts their children.



```{r}
ggplot(data = babies) + 
  geom_jitter(mapping = aes(x = mpregwt, y = bwtoz, color = smoke))+
  geom_smooth(mapping = aes(x = mpregwt, y = bwtoz, color = smoke))
```

#### Our Findings:

* Our team saw that from our plot that women who had a heigher pregnancy weight, had babies with a higher weight which makes sense. Also, we can tell from this plot that babies with mothers who smoked were less likely to have a higher weight and more likely to be a lower weight. Therefore, this could indicate that babies of mothers who smoke are less healthy. 

#### Our Conclusion:

* The data shows that smoking of a mother, leads to a less normal baby weight. This could mean that these babies are less healthy or are more likely to be underweight. 

#### Our Reccomendation: 

* Our reccomendation is that we should continue to analyze more data specifcally trying to learn more about the problem of mothers smoking during or before pregnancy. There should be more legislation put in place to help us to combat the negative impact of smoking especially in regards to mothers. 

## Individual Plots 

#### Kelsey's PLot

```{r}


ggplot(data = babies) + 
  geom_point(mapping = aes(x = number, y = bwtoz, color = smoke))+
  xlab("Number of Cigarettes Smoked") + 
  ylab("Weight of Baby")

```

Kelsey's Findings: I found that there is very little relationship between how many cigarettes smoked a day and a babies birthweight. I found that there are more higher weighted babies as outliers if the mother doesn't smoke.


#### James' Plot 
```{r}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
ggplot(data = babies) +
geom_bar(mapping = aes(x = inc, fill = as.factor(smoke)), position = position_dodge()) 

```

James's Findings:



#### Matt's PLot

```{r}
ggplot(data = cdc, mapping = aes(x = genhlth, y = weight, color=genhlth)) + 
  geom_boxplot() +
  geom_boxplot() +
  coord_flip()
```

Matt's Findings: 

#### Shreeya's PLot

```{r}
ggplot(data = cdc, mapping = aes(x = genhlth, y = weight, color=genhlth)) + 
  geom_boxplot() +
  geom_boxplot() +
  coord_flip()
```

Shreeya's Findings: 

### Team Summary
* I, Kelsey McKenna, created a plot that showed the relationship between the number of cigarettes smoked per day and the child's birthweight. I thought this was an important question and I had predicted that the more a mother smoked, the worse her health and the lower the birthweight of the child. It is clearly not the case and there is little relationship between the two. I used the geom_point function. I added a color function with the variable smoke to highlight if the mother said she smoked or not. I found that there are more higher weighted babies as outliers if the mother doesn't smoke. I added labels to make the plot easier to read.