---
title: "Lab4_LilRascals"
authors: "Matthew Copeland, Kelsey McKenna, James Burke, Shreeya Basrur"
date: "February 7, 2019"
output: html_document
---

```{r}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`) #renaming `bwt/oz` because the "/" sign requires special care
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### Questions/Importance:
# 1) How does smoking affect the number of premature babies are born
This question is very important because it can point to whether or not smoking affects newborns, and if it does, how drastically to affects them. If our findings can give strong evidence toward smoking causing an increased chance of a premature baby, then we could argue that smoking puts the life of your baby at a much higher risk of death. Then we can also use it to advise against smoking during pregnancy.

# 2) What is the relationship between babies weight and whether the mother smokes accounting for a mother's weight during the pregnancy. 
This is an interesting question because the answer could help us to find a better understand of how smoking during pregnancy impacts one's child. This is important because knowing this may help to discourage people from continuing to smoke or starting smoking, and encourage those who want to improve their babies well-being. Especially, in regards to mothers and how it impacts their children.



#### Team Plots:
```{r}
ggplot(data = babies) +
    geom_bar(mapping=aes(x=number, fill = as.factor(Premature)),position = "dodge") + 
    ggtitle("Number of Births vs. Number of Cigarettes Smoked Per Day") +
    xlab("Cigarettes Smoked per Day") +
    ylab("Number of Births")+
    scale_fill_discrete(name = "             Premature births 
(0 = not premature, 1 = premature)")
```



```{r}
ggplot(data = babies) + 
  geom_jitter(mapping = aes(x = mpregwt, y = bwtoz, color = smoke))+
  geom_smooth(mapping = aes(x = mpregwt, y = bwtoz, color = smoke))
```

#### Our Findings:
1) Our findings were that generally the population of women who smoked the most had the fewest babies, and that the proportion of premature to mature born babies was roughly the same for women who didn't smoke and women who smoked multiple cigarettes a day historically. 

2) Our team saw that from our plot that women who had a heigher pregnancy weight, had babies with a higher weight which makes sense. Also, we can tell from this plot that babies with mothers who smoked were less likely to have a higher weight and more likely to be a lower weight. Therefore, this could indicate that babies of mothers who smoke are generally less healthy. 

#### Our Conclusions:
1) The data shows that women who smoked more cigarettes did have a larger proportion of premature babies, although women who smoked more in general had fewer babies overall.  

2) The data shows that smoking of a mother, leads to a less normal baby weight. This could mean that these babies are less healthy or are more likely to be underweight. 

#### Our Reccomendations: 
1) Our recommendation is that further study and analysis is required to produce a more confident conclusion on whether smoking during a pregnancy affects if a baby is born prematurely. However, smoking is definitively injurious to health and this data does show an increased proportion of premature babies born to women who have historically smoked several cigarettes. In general, women should be advised that smoking can affect the birth of a baby and should plan accordingly while further study continues. 

2) Our reccomendation is that we should continue to analyze more data specifcally trying to learn more about the problem of mothers smoking during or before pregnancy. There should be more legislation put in place to help us to combat the negative impact of smoking especially in regards to mothers. 

#### Individual Plots:

## Kelsey's PLot

```{r}


ggplot(data = babies) + 
  geom_point(mapping = aes(x = number, y = bwtoz, color = smoke))+
  xlab("Number of Cigarettes Smoked") + 
  ylab("Weight of Baby")

```
Kelsey's Question: How does whether a mother smokes affect her child's birthweight?

Kelsey's Findings: I found that there is very little relationship between how many cigarettes smoked a day and a babies birthweight. I found that there are more higher weighted babies as outliers if the mother doesn't smoke.


## James' Plot 
```{r}
library(tidyverse)
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
ggplot(data = babies) +
  geom_bar(mapping = aes(x = inc, fill = as.factor(smoke)), position = position_dodge()) +
  xlab("Income")+
  scale_fill_discrete(name = "    Smoking Status
(0 = never, 1 = smokes,
NA = smoked before)")

```
James' Question: Does income affect the proportion of those who smoke, and those who do not smoke?

James' Findings: I found that for the most part a families income does not affect whether or not they smoke. For most income groups the proportion of people who never smoke is greater than that of the proportion of people who smoke.


## Matt's PLot

```{r}
babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`) #renaming `bwt/oz` because the "/" sign requires special care
ggplot(data = babies)+
  geom_jitter(aes(x = parity, y = bwtoz, color = as.factor(marital)),size = 2, na.rm = TRUE)+
  scale_fill_discrete(name = "marrige staus")+
  xlab("Number of Previous Pregnacies")+
  ylab("Weight of Baby at Birth")+
  guides(color=guide_legend("Marrige Status"))
```
Matt's Question: Does a mother having more previous babies affect the other babies after that first one?

Matt's Findings: After looking at the data, babies are born at an average weight by the mothers first couple but as the mother has more babies then the data scatters out and becomes unpredictable. 


## Shreeya's PLot

```{r}
ggplot(data = babies) +
  geom_density(mapping=aes(x=mrace), fill = "yellow", linetype="dashed", size=1) +
  geom_vline(aes(xintercept=mean(mrace)))+
  ggtitle("Density of Race Among Mothers") +
    xlab("Mother's Race(1-5 is white, 6 = mexican, 7 = black, 8 = asian, 9 = mixed)") +
    ylab("Density")
```
Shreeya's Question: What is the distribution of race in the data?

Shreeya's Findings: The vertical black line represents the mean of the race distribution among mothers. This graph shows that a large portion of the participants in this study were white, which could be a confounding variable for the main question. 


#### Team Summary

* I, Kelsey McKenna, created a plot that showed the relationship between the number of cigarettes smoked per day and the child's birthweight. I thought this was an important question and I had predicted that the more a mother smoked, the worse her health and the lower the birthweight of the child. It is clearly not the case and there is little relationship between the two. I used the geom_point function. I added a color function with the variable smoke to highlight if the mother said she smoked or not. I found that there are more higher weighted babies as outliers if the mother doesn't smoke. I added labels to make the plot easier to read. I also created the team plot number 1 and the question and analysis for our first team plot. 

* I, Shreeya Basrur,created a density plot to analyse the distribution of race demographics for mothers in this study. I filled in the density plot yellow for visibility, and added the mean line to clarify the general density plot. I also added axis titles as well as an informative title to be easier to read. This plot was created with the geom_density function with aesthetic modifiers that filled in the plot yellow and added the dashed line. Geom_vline was the function I used to create the mean line for this plot. I also created team plot two in order to answer our second question. 

* I, James Burke, wrote the first question (smoking vs premature babies) and its importance in warning mothers of the affect smoking can have on their newborns. I also helped to create and format the Number of Births vs. Number of Cigarettes Smoked Per Day plot. For my individual plot I created a bar chart that compares the proportion of smokers vs non smokers, to they income grouping. I made incomne the x axis and then had separate bars in each income grouping, for peoples smoking status. Having these bars side by side makes it easy to visualize the proportion. I did this using the fill aestetic and by making the smokers variable a factor, then I used position dodge to separate the bars. I also changed the x axis label along with the legend title.

* I, Matthew Copeland, created a graph looking at the effects of having more babies on the baby more than the mother. This can help show that the toll of giving birth can affect the mother but can affect previous babies as well. I created a scatter plot graph where the x-axis is the number of previous pregnancies and the y-axis to be weight of baby at birth. The color legend is based of the marital status of the family.